The states of the controller FSM are:
\begin{itemize}
  \item rst: Reset
  \item clr: Clear the counter, memory and input FSM.
  \item max: The maximal number has been reached.
  \item maxb (max\_bar): The maximal number has \textbf{not} been reached
  \item cnt\_upg (count up in green phase): The counter counts up because of the input \textit{action\_i(0)}. This happens with a green LED on if the maximal number of persons in the room has not been reached (\textit{cnt\_i} < max).
  \item en\_reggu (enable register write in green phase while up-count): The memory for the counters value needs to be registered.
  \item st\_uartgu (start UART-TX in green phase while up-count): The UART starts sending the counter value.
  \item st\_s3gu (start S3-IF in green phase while up-count): The IF starts sending the counter value and the respective action (person entered room).
  \item cnt\_dng (count down in green phase): The counter counts down because of the input \textit{action\_i(1)}. This happens with a green LED on if the maximal number of persons in the room has not been reached (\textit{cnt\_i} < max).
   \item en\_reggd (enable register write in green phase while down-count): The memory for the counters value needs to be registered.
  \item st\_uartgd (start UART-TX in green phase while down-count): The UART starts sending the counter value.
  \item st\_s3gd (start S3-IF in green phase while down-count): The IF starts sending the counter value and the respective action (person left room).
  \item cnt\_upr (count up in red phase): The counter counts up because of the input \textit{action\_i(0)}. This happens with a red LED on if the maximal number of persons in the room has been reached (\textit{cnt\_i} >= max).
  \item en\_regru (enable register write in red phase while up-count): The memory for the counters value needs to be registered.
  \item st\_uartru (start UART-TX in red phase while up-count): The UART starts sending the counter value.
  \item st\_s3ru (start S3-IF in red phase while up-count): The IF starts sending the counter value and the respective action (person entered room).
  \item cnt\_dnr (count down in red phase): The counter counts down because of the input \textit{action\_i(1)}. This happens with a red LED on if the maximal number of persons in the room has been reached (\textit{cnt\_i} >= max).
   \item en\_regrd (enable register write in red phase while down-count): The memory for the counters value needs to be registered.
  \item st\_uartrd (start UART-TX in red phase while down-count): The UART starts sending the counter value.
  \item st\_s3rd (start S3-IF in red phase while down-count): The IF starts sending the counter value and the respective action (person left room).
  \item en\_reggu (enable register in green phase while up-count):
\end{itemize}
